% This file is part of the CTAN package named gfnotation.
%
%   This is the users guide to the TeX macros ``GFnotation.tex''.
%   GFnotation.tex: typeset G. Frege's notation of the Begriffsschrift or
%                   the Grundgesetze der Arithmetik in plain TeX
%
%   Copyright (C) 2015  Udo Wermuth (author)
%
%   This program is free software: you can redistribute it and/or modify
%   it under the terms of the GNU General Public License as published by
%   the Free Software Foundation, either version 3 of the License, or
%   (at your option) any later version.
%
%   This program is distributed in the hope that it will be useful,
%   but WITHOUT ANY WARRANTY; without even the implied warranty of
%   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%   GNU General Public License for more details.
%
%   You should have received a copy of the GNU General Public License
%   along with this program.  If not, see <http://www.gnu.org/licenses/>.
% %%%
% %%% GFnotation
% %%%
\input amssym.def
\catcode`\ß=11
\let\nosubst=f
\let\showrotation=t \input rotate
\let\fgefontsknown=t
\input GFnotation
\gfbslognotationtrue
\gfbszeigestatstrue
% %%%
% %%% verbatim macros (from manmac.tex)
% %%%
\newskip\ttglue {\tt\global\ttglue=0.5em plus 0.25em minus 0.15em }
\def\ttverbatim{\begingroup \frenchspacing
  \catcode`\\=12 \catcode`\{=12 \catcode`\}=12 \catcode`\$=12 \catcode`\&=12
  \catcode`\#=12 \catcode`\%=12 \catcode`\~=12 \catcode`\_=12 \catcode`\^=12
  \obeyspaces \obeylines \tt}
\def\verbatimspace{\ifvmode\indent\fi\space}
{\obeyspaces \gdef\makespaceverbspace{\def {\verbatimspace}}}
\outer\def\verbatim{$$\ifdim\parskip>0pt
    \abovedisplayskip=\parskip \abovedisplayshortskip=\parskip
    \belowdisplayskip=\parskip \belowdisplayshortskip=\parskip
  \else
    \abovedisplayskip=3pt \abovedisplayshortskip=3pt
    \belowdisplayskip=3pt \belowdisplayshortskip=3pt
  \fi
  \let\par=\endgraf \ttverbatim \makespaceverbspace \parskip=0pt
  \catcode`\§=0 \advance\leftskip by 10pt \ttfinish}
{\catcode`\§=0 §catcode`§\=12 % § is temporary escape character
  §obeylines % end of line is active
  §gdef§ttfinish#1^^M#2\endverbatim{§vbox{#2}§endgroup$$}}
\catcode`\¤=\active
{\obeylines \gdef¤{\ttverbatim \spaceskip\ttglue \let^^M=\  \let¤=\endgroup}}
% %%%
% %%% Examples
% %%%
\def\Econt{\goodbreak\bgroup
  \parskip=0pt \leftskip=3\parindent \rightskip=\leftskip \parindent=0pt }
\def\Epause{\par\egroup}
\def\Example#1{\bigskip\Econt\noindent{\bf Example #1.}\ }
\def\Eend#1{\ifnum#1>0 \vskip-\baselineskip\fi
  \hfill\vrule height 7pt depth 1pt width 3pt \Epause\bigskip}
% %%%
% %%% hyphenation exceptions
% %%%
\hyphenation{Gottlob Frege Begriffs-schrift Grund-gesetze}
% %%%
% %%% other
% %%%
\def\\{\par\vskip0pt plus.3\vsize\penalty-253\vskip0pt plus-.3\vsize\relax}
\def\AmSTeX{$\cal A\kern-.1667em\lower.5ex\hbox{$\cal M$}\kern-.075em S$-\TeX}
\def\La{L\kern-3pt \raise.42ex\hbox{$\scriptstyle A$}}
\def\LaTeX{\La\kern-.15em\TeX}
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\parskip=0.25\baselineskip
\settabs 6 \columns
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\font\titlefont=cmssdc10 at 36pt
\font\subtitlefont=cmssdc10 at 17pt
%
\centerline{\titlefont Users Guide}
\bigskip
\centerline{\subtitlefont GFnotation.tex (version 2.9, 29.03.2015)}

% %%%%%%%%%%%
\beginsection 1. Description

This package implements macros to typeset formulas and inferences in
Gottlob Frege's notation in the style of his ``Begriffsschrift, eine der
arithmetischen nachgebildete Formelsprache des reinen Denkens'' (see
[1]) or in the style of his main work ``Grundgesetze der Arithmetik''
(see [2]). Here is an example of a single
formula: \frege|{..\gA.{*.a.{f(\da)}}}. And this is a complete
inference:
\nlp4\*_[-\*-:-\ce{$\gA$}
\nlp4\*___\*_'-\ce{$\gB$}
\nlp4\*_[-\*---\ce{$\gB$}
\nlp4\*---\*---\*---
\nlp4\*_[-\*---\ce{$\gA$} .

The package has two input formats, the so-called ``symbolic
representation'' and the ``short form.''  The short form outputs the
formulas either in the style of Frege's {\it Begriffsschrift\/} or in
the style that is used in his {\it Grundgesetze}.  The above formulas
are in the style of the {\it Begriffsschrift\/}; here is the first one
in the style of the {\it Grundgesetze\/}:
\toggleGGstyle\frege|{..\gA.{*.a.{f(\da)}}}.\toggleGGstyle

{\it Note\/}: This users guide shows how to work with the package.
It does not explain the constructions and symbols in Frege's
notation.


% %%%%%%%%%%%
\beginsection 2. Elements of Frege's notation

{\bf a) Letters.}\ Frege used lowercase roman letters, $a$, $b$,
$c$, \dots, uppercase Greek letters $\gA$, $\gB$, $\gG$, \dots,
lowercase Greek letters $\ka$, $\kb$, $\kg$, \dots, and Fraktur
letters $\da$, $\de$, $\dF$, \dots\thinspace.

\noindent
{\bf b) Formulas (Begriffsschrift).}\
The following notation is used by Frege to build formulas:
\smallskip
\+&{\it Content stroke\/}: \*---\ze{$\gA$}.
 &&{\it Judgment\/}: \*[--\ze{$\gA$}.\cr
\smallskip
\+&{\it Negation\/}: \*-~-\ze{$\gA$}.
 &&{\it Affirmation\/}: \*-+-\ze{$\gA$}.\cr
 \smallskip
\+&{\it Condition\/}: \frege .{..\gB.\gA}.
 &&{\it Generality\/}: \frege .{*.a.{\gF(\da)}}.\cr
\+&{\it Condition stroke\/}: \*_!_.
 &&{\it Definition\/}: \*"3- $((\gA)\equiv\gB)$.\cr
\smallskip
\+&{\it Substitution\/}: $a$\*_|_$f(a)$.\cr

\noindent
{\bf c) Inferences (Begriffsschrift).}\
For inferences two types of lines are used:
\smallskip
\+&{\it Single line\/}: \*---\*---.
 &&{\it Double lines\/}: \*===\*===.\cr
\smallskip
\+&{\it Separator between inference chains\/}:\cr
\decollator4

\noindent
{\bf d) Terminal symbols (Begriffsschrift).}\
Four special constructions appear in the {\it Begriffsschrift\/}:
\smallskip
\+&$\gfvererbe\kd\ka{{x(\ka)}}{{f(\kd,\ka)}}$, \quad$\gffolgt\kg\kb{f(x_\kg,y_\kb)}$,
\quad$\gfgehoertan\kg\kb{f(x_\kg,z_\kb)}$, \quad$\gfeindeutig\kd\ke{f(\kd,\ke)}$.\cr

\toggleGGstyle
\gfbskeinekompaktehoehlungtrue
\noindent
{\bf e) Formulas (Grundgesetze).}\
The following notation is used by Frege to build formulas:
\smallskip
\+&{\it Content stroke\/}: \frege.{=.\gA}.
 &&{\it Judgment\/}: \frege|{=.\gA}.\cr
\smallskip
\+&{\it Negation\/}: \frege.{=-\gA}.
 &&{\it Affirmation\/}: \frege.{=+\gA}.\cr
 \smallskip
\+&{\it Condition\/}: \frege .{..\gB.\gA}.
 &&{\it Generality\/}: \frege .{*.a.{\gF(\da)}}.\cr
\+&{\it Condition stroke\/}: \*_!_.
 &&{\it Definition\/}: \*"39 $\gA=\gB$.\cr

\noindent
{\bf f) Inferences (Grundgesetze).}\
For inferences eight types are used:
\smallskip
\+&{\it Single line\/}: \*?--\*?--.
 &&{\it Double lines\/}: \*?==\*?==.\cr
\smallskip
\+&{\it Dashed line\/}: \*?-_\*?-_.
 &&{\it Dashed double lines\/}: \*?=_\*?=_.\cr
\smallskip
\+&{\it Single/double line\/}: \*?=*\*?=*.
 &&{\it Dot/dashed line\/}: \*?.-\*?.-.\cr
\smallskip
\+&{\it Contraposition\/}: \*?>x.
 &&{\it Quantification\/}: \*?>u.\cr
\smallskip
\+&{\it Separator\/}: \gfggtrenner.\cr

\noindent
{\bf g) Special symbols (Grundgesetze).}\
Several special characters appear in the {\it Grundgesetze\/}:
\smallskip
\+&{\it symbols\/}: $\Gbf$ \G0 \G1 $\Guu$\cr
\smallskip
\+&{\it unary operators\/}: \G! $\kasl$ \GI\ $\G>$ \G- \G~ $\G*$
                            $\G6$ $\Gf$ $\Gp$ $\GUorg$ $\Gnorg$
&&&{\it last two from the fge fonts\/}: $\GU$ $\Gn$\cr
\smallskip                            
\+&{\it binary operators\/}: $\G^$ $\G_$ $\G;$ $\Gfm$ $\Gsm$
                             $\G<$ $\Gmf$ $\Gl$
&&&{\it rotated\/}: $\Gy$ $\Ge$ $\Gc$ $\GF$ $\Gss$\cr
\toggleGGstyle


% %%%%%%%%%%%
\beginsection 3. How to load the package

The package requires characters that come with \AmSTeX\ (see [3]);
the following statement must be given before the package is loaded:
¤\input amssym.def¤.  Then the package is loaded with the statement
¤\input GFnotation¤.

The package has two styles: the one that is used in the {\it
Begriffsschrift\/} and the one that is used in the {\it Grundgesetze}.
If you want to work with the style of the {\it Grundgesetze\/} several
external flags are available.

First, if only inferences without substitutions are needed the line
¤\let\nosubst=t¤ should be entered before ¤GFnotation¤ is loaded. This
avoids typesetting errors and saves some \TeX\ registers.  Second, the
flag ¤\let\showrotation=t¤ can be set to output some terminal symbols
of the {\it Grundgesetze\/} rotated. Then ¤\input rotate¤ (see [4])
must be stated too.  Otherwise the symbols are not shown as rotated
symbols (some DVI screen drivers can't do PS-rotation). The affected
symbols are shown in Section~2g.  Third, the flag
¤\let\fgefontsknown=t¤ can be set to indicate that the fge-fonts by
J.~J.~Green (see~[5]) are available. Then some symbols of these fonts
are used in the output. The affected symbols are shown in Section~2g.

After the package has been loaded two flags can be set (independent of the
style).
\item{1.}¤\gfbslognotationtrue¤ writes all formulas, which are given in the
short form, in their symbolic representation into the log file of the
\TeX\ run.
\item{2.}¤\gfbszeigestatstrue¤ writes a \TeX\ message for each formula in
short form if its number of lines is a new maximum.

By default the package activates the style that is used in the {\it
Begriffsschrift}. To switch to the style of the {\it Grundgesetze\/}
the global acting command ¤\toggleGGstyle¤ shall be given.

Therefore a typical start is coded in the following way if the style
of the {\it Begriffschrift\/} is required:
\verbatim
\input amssym.def
\input GFnotation
\gfbslognotationtrue % log short form formulas in symbolic representation
\gfbszeigestatstrue  % inform about maximum number of lines of short form formulas
\endverbatim

Otherwise, the style of the {\it Grundgesetze\/} is started with
\verbatim
\input amssym.def
%%% external flags:
\let\nosubst=t       % substitutions aren't needed
\let\showrotation=t \input rotate % activate this line if PS-rotation works
%\let\fgefontsknown=t % activate this line if the fge-fonts are available
\input GFnotation
%%% internal flags:
\gfbslognotationtrue % log short form formulas in symbolic representation
\gfbszeigestatstrue  % inform about maximum number of lines of short form formulas
%%% define the style of the output:
\toggleGGstyle       % use the style of the Grundgesetze
\endverbatim


% %%%%%%%%%%%
\beginsection 4. Shortcuts for letters

The characters that Frege used (except the roman letters) are coded
with two letter shortcuts.  The first letter indicates how the second
letter is output.  Uppercase Greek letters are introduced by `g',
lowercase Greek letters by `k', and the Fraktur letters by `d':
¤$\gA$¤ outputs~$\gA$, ¤$\ka$¤ outputs~$\ka$, and ¤$\da$¤
outputs~$\da$.

Here is the list of the uppercase Greek letters; the coding follows~[6].
\smallskip
\+&¤$\gA$¤: $\gA$&  ¤$\gB$¤: $\gB$&  ¤$\gG$¤: $\gG$&  ¤$\gD$¤: $\gD$&  ¤$\gE$¤: $\gE$\cr
\+&¤$\gZ$¤: $\gZ$&  ¤$\gH$¤: $\gH$&  ¤$\gJ$¤: $\gJ$&  ¤$\gI$¤: $\gI$&  ¤$\gK$¤: $\gK$\cr
\+&¤$\gL$¤: $\gL$&  ¤$\gM$¤: $\gM$&  ¤$\gN$¤: $\gN$&  ¤$\gX$¤: $\gX$&  ¤$\gO$¤: $\gO$\cr
\+&¤$\gP$¤: $\gP$&  ¤$\gR$¤: $\gR$&  ¤$\gS$¤: $\gS$&  ¤$\gT$¤: $\gT$&  ¤$\gU$¤: $\gU$\cr
\+&¤$\gF$¤: $\gF$&  ¤$\gQ$¤: $\gQ$&  ¤$\gY$¤: $\gY$&  ¤$\gW$¤: $\gW$\cr

For the Fraktur letters only a subset of the alphabets is provided;
these are the letters that appear in the {\it Begriffsschrift\/} or in
part~1 of Vol.~1 of the {\it Grundgesetze\/}:
\smallskip
\+&¤$\da$¤: $\da$&  ¤$\dc$¤: $\dc$&  ¤$\dd$¤: $\dd$&  ¤$\de$¤: $\de$&  ¤$\df$¤: $\df$\cr
\+&¤$\dg$¤: $\dg$&  ¤$\dm$¤: $\dm$&  ¤$\dn$¤: $\dn$&  ¤$\ddo$¤: $\ddo$&¤$\ddp$¤: $\ddp$\cr
\+&¤$\dq$¤: $\dq$&  ¤$\dr$¤: $\dr$&  ¤$\dt$¤: $\dt$&  ¤$\du$¤: $\du$&  ¤$\dF$¤: $\dF$\cr

{\it Note\/}: Commands like ¤\dp¤ or ¤\do¤ shouldn't change their
plain \TeX\ meaning. Therefore the commands for the letters `o' and
`p' are build with two `d's.

\begingroup\def\ds{{\frak s}} % keep this definition local
You can add more letters using a simple
¤\def¤. For example, to add $\ds$ declare ¤\def\ds{{\frak s}}¤.
\endgroup


% %%%%%%%%%%%
\beginsection 5. Input of formulas using the
                 ``symbolic representation'' (Begriffsschrift)

This input format gives complete control over the output of the
formulas.  But it requires some amount of typing.  All formulas are
built from symbols and three of them are grouped into the argument of
a control symbol.  The control symbol is the ¤\*¤.  (The original
meaning as discretionary multiplication is kept in the new control
word ¤\discretionarytimes¤.)

In total there are 35 symbols. In the following list the parentheses
contain the sequence ¤\*"#"¤ if symbol `{\tt\#}' is explained.  The
first collection lists the important symbols that are used for
hand-coding the formulas in Frege's notation.
\smallskip
% 11 symbols
\item{\tt\char95 }(\*"_") is the empty space;
\item{\tt-}(\*"-") is a horizontal line for the content stroke and the inference rule;
\item{\tt\~}(\*"~") is a horizontal line with centered negation;
\item{\tt+}(\*"+") is a horizontal line with two centered negation indicators (affirmation);
\item{\tt:}(\*":") is the then-part of the condition;
\item{\tt'}(\*"'") is the if-part of the condition;
\item{\tt!}(\*"!") is a vertical line for the condition stroke;
\item{\tt[}(\*"[") is a vertical line placed left and starts a horizontal line;
\item{\tt|}(\*"|") is the vertical line for the substitution;
\item{\tt3}(\*"3") are two vertical lines used for {\it definitions\/};
\item{\tt=}(\*"=") are two horizontal lines.

\smallskip\noindent
The following symbols extend the basic list and can be used to
fine-tune the output.
\smallskip
% 8 symbols (sum=19)
\item{\tt"}(\*""") is a skip, that is, no output at all;
\item{\tt(}(\*"(") negation but the indication of negation is moved to the left;
\item{\tt)}(\*")") negation with the indicator moved to the right;
\item{\tt<}(\*"<") negation: the indicator is placed at the left end;
\item{\tt>}(\*">") negation: the indicator is placed at the right end;
\item{\tt\^{}}(\*"^") affirmation with indicators moved to the left;
\item{\tt/}(\*"/") affirmation with indicators moved to the right;
\item{\tt]}(\*"]") is a vertical line placed right, that is, only the judgment stroke is printed.

For the generality a letter must be given. Therefore the parentheses
contain the sequence ¤\*#a.¤ (¤\*@a"¤ for {\tt @}). Instead of the
letter `a' any other letter can be used that is available in the
Fraktur font (see Section~4).  A triple for the generality must be
started with one of the following symbols {\tt.,`;@}.  If it is not
the {\tt @} then the symbol after the letter must be one of
{\tt.,`;}\thinspace.
\smallskip
% 5 symbols (sum=24)
\item{\tt.}(\*.a.) is a short vertical line;
\item{\tt,}(\*,a.) is a short vertical line with a centered negation indicator;
\item{\tt`}(\*`a.) represents negation but the indicator is moved to the left;
\item{\tt;}(\*;a.) is a short vertical line with affirmation;
\item{\tt @}(\*@a") signals (without any output) that a letter for the generality follows.

{\it Note\/}: The letter `{\tt\ss}' is special. It uses the content of
the token register ¤\dsz¤. For example, with the default ¤\dsz={\xi}¤
¤\*.¤{\tt\ss}¤.¤ outputs \*.ß.. Of course, {\tt\ss} must be a letter
so for plain \TeX\ the command ¤\catcode`\¤{\tt\ss}¤=11¤ must be
entered before ¤\input GFnotation¤ is done.

\Example1
The formula \
\vtop{%
  \hbox{\strut\*[-:\*--- $\gA$}
  \hbox{\strut\*__'\*.a. $f(\da)$}}
of Section~1 can be coded in the following way:
\verbatim
\vtop{%
  \hbox{\strut\*[-:\*--- $\gA$}
  \hbox{\strut\*__'\*.a. $f(\da)$}}
\endverbatim\Eend1

The last list contains symbols that are used by the short form. The
symbols do not try to have a mnemonic representation for their
output. Most of these symbols can ``change'' and stand for one of two
symbols from the above lists.
\smallskip
% 11 symbols (sum=35)
\item{\tt0}negation either moved to the left or the right: {\tt0=()}
\item{\tt1}affirmation: {\tt1=\^{}/}
\item{\tt2}again affirmation: {\tt2=\^{}+}
\item{\tt4}not required content strokes: {\tt4=-"}
\item{\tt5}not required empty space: {\tt5=\char95"}
\item{\tt6}judgment strokes: {\tt6=[]}
\item{\tt7}negated generality: {\tt7=`,}
\item{\tt8}again negation: {\tt8=<\~{}}
\item{\tt9}represents a content stroke that has only one third of its usual length;
\item{\tt*}is a skip whose length is multiplied by $5/3$ to get back into alignment after a {\tt9}; 
\item{\tt\$} is used only with a generality and represents either a short content stroke or an empty space.


% %%%%%%%%%%%
\beginsection 6. Input for inference chains in the
                 ``symbolic representation'' (Begriffsschrift)

In order to typeset complete inferences several formulas must be
positioned in a chain and the rule for the inference has to be entered.
The construction of Example~1 with ¤\vtop¤, ¤\hbox¤, and ¤\strut¤ is
shortened to a single command ¤\nlp¤. The terminal symbols become
an argument of a second command ¤\ce¤.
\smallskip
\item{1.}¤\nlp<n>¤ starts a new line and sets the following text {\tt<n>} units to the right;
\item{2.}¤\ce{<s>}¤ outputs the string {\tt<s>} with some spacing at the end of a line.

\Example2
This is the implementation of the inference in Section~1, that is
\nlp4\*_[-\*-:-\ce{$\gA$}
\nlp4\*___\*_'-\ce{$\gB$}
\nlp4\*_[-\*---\ce{$\gB$}
\nlp4\*---\*---\*---
\nlp4\*_[-\*---\ce{$\gA$}\qquad:
\smallskip
\verbatim
\nlp4\*_[-\*-:-\ce{$\gA$}
\nlp4\*___\*_'-\ce{$\gB$}
\nlp4\*_[-\*---\ce{$\gB$}
\nlp4\*---\*---\*---
\nlp4\*_[-\*---\ce{$\gA$} .
\endverbatim\Eend1

\noindent
For complete inferences with substitutions more commands are
required:
\smallskip
\item{3.}¤\ci{<s>}¤ is like ¤\ce¤ but must be used in the line and spaces that follow are ignored;
\item{4.}¤\nlpc<n>{<s>}¤ acts like ¤\nlp<n>¤ and sets the string {\tt<s>} to the left in the white space;
\item{5.}¤\fono{<m>}¤ outputs {\tt<m>} as the number of the formula at the right end of the previous line;
\item{6.}¤\rep<n><f>¤ repeats the ¤\*¤-formula {\tt<f>} {\tt<n>} times;
\item{7.}¤\bcc<n>/<p><m>:<c><f>\ecc¤ typesets an inference rule.
{\tt<n>} is the position to the right, {\tt<m>} is a formula number used in the inference,
and {\tt<f>} is the rule itself. The parameter {\tt<p>} is either `l', 'm', or `r' to
position the formula number to the left, in the middle, or to the right.
And {\tt<c>} is either `.'\ or `:'\ to output one or two colons after the formula number.

\Example3
The following code
\Epause
\verbatim
     \nlp4\ci{\hss\thinspace58\hss}\rep3\*___\*_[-\*-:-\*---\*-:-\ce{$f(b)$}        
\nlpc4{$f(A)$}\*_|_\*-:-\ci{$f(A)$}\rep2\*___\*_!_\*___\*_'-\ce{$g(b)$}        
         \nlp4\*_|_\*_'-\ci{$g(A)$}\rep2\*___\*_'-\*.a.\*-:-\ce{$f(\da)$}      
   \nlpc4{$c$}\*_|_\ci{$b$}        \rep3\*___\*___\*___\*_'-\ce{$g(\da)$}      
        \bcc4/m30:.\rep3\*___\null      \rep6\*---\ecc
   \nlpc4{$a$}\*_|_\ci{$f(b)$}     \rep2\*___\*_[-\*-:-\*-:-\*,a.\*-:-\ce{$f(\da)$} 
   \nlpc4{$c$}\*_|_\ci{$g(b)$}     \rep2\*___\*___\*_!_\*_!_\*___\*_'-\ce{$g(\da)$} 
   \nlpc4{$b$}\*_|_\*.a.\*-:-\ci{$f(\da)$}   \*___\*_!_\*_'-\*---\*--(\ce{$f(b)$}   
         \nlp4\*_|_\*___\*_'-\ci{$g(\da)$}   \*___\*_'-\*---\*---\*---\ce{$g(b)$}   
\fono{59}
\endverbatim\Econt

outputs formula~59 of the {\it Begriffsschrift}:
\smallskip
     \nlp4\ci{\hss\thinspace58\hss}\rep3\*___\*_[-\*-:-\*---\*-:-\ce{$f(b)$}        
\nlpc4{$f(A)$}\*_|_\*-:-\ci{$f(A)$}\rep2\*___\*_!_\*___\*_'-\ce{$g(b)$}        
         \nlp4\*_|_\*_'-\ci{$g(A)$}\rep2\*___\*_'-\*.a.\*-:-\ce{$f(\da)$}      
   \nlpc4{$c$}\*_|_\ci{$b$}        \rep3\*___\*___\*___\*_'-\ce{$g(\da)$}      
        \bcc4/m30:.\rep3\*___\null      \rep6\*---\ecc
   \nlpc4{$a$}\*_|_\ci{$f(b)$}     \rep2\*___\*_[-\*-:-\*-:-\*,a.\*-:-\ce{$f(\da)$} 
   \nlpc4{$c$}\*_|_\ci{$g(b)$}     \rep2\*___\*___\*_!_\*_!_\*___\*_'-\ce{$g(\da)$} 
   \nlpc4{$b$}\*_|_\*.a.\*-:-\ci{$f(\da)$}   \*___\*_!_\*_'-\*---\*--(\ce{$f(b)$}   
         \nlp4\*_|_\*___\*_'-\ci{$g(\da)$}   \*___\*_'-\*---\*---\*---\ce{$g(b)$}   
\fono{59}\Eend0

\noindent
One more command is required for collections of inferences:
\item{8.}¤\decollator<l>¤ outputs a thick line of width ¤0.<l>\hsize¤. The line is
centered. It is preceded and followed by vertical space.

\Example4
Here are some inferences of the {\it Begriffsschrift\/}:
\outof p3,0"1"with\thatis
\formula p9|{..{..a.b}.a}
\followswith p3"8"a.p8s5
\substituting p3 d:a \whichgives
\formula p9|{..{..a.a}.b}
\named "26"
\decollator4
\outof p3,0"26"with b:{..{..a.b}.a} \thatis
\formula p8|{..{=.{..a.a}}.{..{..a.b}.a}}
\followswith p3"1"a:p8s5
\formula p8|{=.{=.{..a.a}}}
\named "27"

The command ¤\decollator4¤ appears after formula~26.
\Eend0


% %%%%%%%%%%%
\beginsection 7. Input of formulas using the ``short form'' 

This input format reduces the amount of typing compared to the
symbolic representation.  It follows given rules and produces as an
intermediate step the code of the symbolic representation.

The input is recursively entered using four constructions:
\smallskip
\item{1.}{\tt=<}sign{\tt>\char"7B<}formula{\tt>\char"7D} to add a
signed content stroke in front of a formula;
\item{2.}{\tt<}sign{\tt><}sign{\tt>\char"7B<}formula{\tt>\char"7D<}%
                          sign{\tt>\char"7B<}formula{\tt>\char"7D}
to build a signed condition from two signed formulas;
\item{3.}{\tt*<}sign{\tt><}character{\tt><}sign{\tt>\char"7B<}%
                           formula{\tt>\char"7D} to add a signed
concavity in front of a signed formula;
\item{4.}{\tt!\char92<}nameofdef{\tt><}arguments{\tt>} to insert a
previously defined subformula.

\noindent
Here {\tt<}formula{\tt>} is either a valid formula constructed with
the four rules or a terminal string (both are typeset in math mode)
and {\tt<}sign{\tt>} is one of the following symbols:
\item{\tt.} represents no sign;
\item{\tt-} represents the negation;
\item{\tt+} represents the affirmation.

In rule~4 {\tt<}nameofdef{\tt>} is together with the preceding
backslash a defined control word that has at most three arguments
given in {\tt<}arguments{\tt>}.  The control word must define a part
of a formula.  This rule implements {\sl macro expansion\/} to avoid
retyping of often used parts in formulas.

Most formulas are created by rules~2 and~3; rule~1 is only needed to
type a single line.

{\it Important note\/}: The braces around {\tt<}formula{\tt>} can be
omitted if the formula is a terminal string that consists of a single
token. Double curly braces are required around terminal strings that
have more than five tokens.  But single braces must always be used for
formulas in rules 1--3, which are not terminal strings.  (In Example~5
below case~i shows that a single token needs no braces.  The cases ii
and~iii show the use of single and double curly braces.)

Formulas that are created by the above rules must be given to the
following commands:
\smallskip
\item{1.}¤\frege<j>{<f>}¤ where {\tt<f>} is a formula created by the above
stated rules and {\tt<j>} is either `{\tt.}'\ or `{\tt|}'.
\item{2.}¤\formula p<n><j>{<f>}¤ adds a position parameter {\tt<p>} to
move the formula {\tt<n>} units to the right.

\noindent
The formula {\tt<f>} is output as a judgment if {\tt<j>} equals
{\tt|}.  The command ¤\frege¤ shall be used in running text,
¤\formula¤ in vertical mode. But ¤\frege¤ doesn't start a paragraph;
use ¤\indent¤ (or ¤\noindent¤) if ¤\frege¤ is the first item
in a paragraph.

\Example5
\def\0#1{*.a.{#1(\da)}}
Here are a few examples for all four rules:
\smallskip
\advance\leftskip by 20pt
\item{i)}¤\frege|{=-\gA}¤ outputs \frege|{=-\gA}.
\item{ii)}¤\frege.{-.\gA-{{g(\ka,\kb)}}}¤ outputs \frege.{-.\gA-{{g(\ka,\kb)}}}.
\item{iii)}¤\frege.{*-a+{f(\da)}}¤ outputs \frege.{*-a+{f(\da)}}.
\item{iv)}¤\def\0#1{*.a.{#1(\da)}}¤ together with ¤\frege|{-.\gA-{!\0f}}¤
outputs \frege|{-.\gA-{!\0f}}.
\smallskip
\advance\leftskip by -20pt
The output of iv) is nearly the formula of Section~1 (\frege|{..\gA.{!\0f}});
it can be coded (without ¤\0¤) as ¤\frege|{..\gA.{*.a.{f(\da)}}}¤.
\Eend0

Of course, a formula in Frege's notation that has several lines is not
easy to read in a sequential form. The periods and braces do not help
to make the structure of that formula easy to recognize.
I suggest to copy a Frege formula from left to right and from bottom
to top and to use more than one line for the sequential coding.

\Example6
This example shows how to construct in a structured manner the short
form of a given formula, which is written in Frege's notation.
The formula is \frege|{.-{..\gA.\gB}.{*.a.{..\gG-\gD}}}.

Step~1: The formula is a judgment and starts at the left with a
condition, whose upper line contains a negation indicator.
So the coding starts with
\verbatim
\frege|{.-{}.{}}
\endverbatim

Or write it in two lines to make the structure visible:
\verbatim
\frege|{.-{}
         .{}}
\endverbatim

Step~2: The bottom part starts with a concavity, so the input becomes
\verbatim
\frege|{.-{}
         .{*.a.{}}}
\endverbatim

Step~3: A simple condition comes after the concavity
\verbatim
\frege|{.-{}
         .{*.a.{..\gG-\gD}}}
\endverbatim

And the final step is to add the condition in the upper part:
\verbatim
\frege|{.-{..\gA.\gB}
         .{*.a.{..\gG-\gD}}}
\endverbatim
\medskip
{\it Note\/}: The input
\verbatim
\frege|{..{-.\gA.\gB}
         .{*.a.{..\gG-\gD}}}
\endverbatim
outputs
\frege|{..{-.\gA.\gB}
         .{*.a.{..\gG-\gD}}},
which looks very similar to the given formula but, of course,
the negation indicator in the upper part has moved a little bit
to the right.
\Eend0


% %%%%%%%%%%%
\beginsection 8. Inferences of the Begriffsschrift in the ``short form'' 

Five commands are available to typeset inferences with the short form:
\smallskip
\item{1.}¤\outof p<n>,<o>"<m>"with<r>\thatis¤ starts the inference chain.
The parameter {\tt<n>} is the number of units the output is moved to
the right, {\tt<m>} is a number of a formula, {\tt<o>} is an offset to
position the number {\tt<m>} {\tt<o>} lines lower, and {\tt<r>} is
either empty or a space separated list of pairs {\tt v:F} to indicate
the substitution of {\tt v} by the formula {\tt F}.
\item{2.}¤\use p<n>"<m>"¤ is a shortcut of ¤\outof¤ if the parameters {\tt<o>}
and {\tt<r>} are not needed; {\tt<o>} is set 0 and {\tt<r>} is empty.
\item{3.}¤\followswith p<n>"<m>"a<c>p<N>s<l>¤ draws the inference rule and
can be expressed in the ¤\bcc¤ macro of Section~6, no~7, as
¤\bcc{<n>}/m{<m>}:<c>\rep(<N>-<n>)\*___\rep{<l>}\*---\ecc¤.
\item{4.}¤\substituting p<n> <r>\whichgives¤ has the usual position parameter
{\tt<n>} and a list of substitutions {\tt<r>} in the above described form {\tt
v:F}.
\item{5.}¤\named "<m>"¤ outputs the number {\tt<m>} as the number of the
previous formula.

An inference chain is typically typed with the sequence:
\verbatim
\outof ... with ... \thatis
\formula ...
\followswith ...
\substituting ... \whichgives
\formula ...
\named ...
\endverbatim

\noindent
or if no substitution in the first formula is required, the ¤\outof¤ line
is replaced by the simpler ¤\use¤.

\Example7
The formula of Example~3
\smallskip
\outof p4,0"58"with f(A):{..{f(A)}.{g(A)}}
                       c:b
\thatis
\formula p8|{..{..{f(b)}.{g(b)}}
              .{*.a.{..{f(\da)}.{g(\da)}}}}
\followswith p4"30"a.p8s6
\substituting p4 a:{f(b)}
                 c:{g(b)}
                 b:{*.a.{..{f(\da)}.{g(\da)}}}
\whichgives
\formula p8|{..{..{*-a.{..{f(\da)}.{g(\da)}}}
                 -{f(b)}}
              .{g(b)}}
\named "59"

\bigskip
has this coding:
\verbatim
\outof p4,0"58"with f(A):{..{f(A)}.{g(A)}}
                       c:b
\thatis
\formula p8|{..{..{f(b)}.{g(b)}}
              .{*.a.{..{f(\da)}.{g(\da)}}}}
\followswith p4"30"a.p8s6
\substituting p4 a:{f(b)}
                 c:{g(b)}
                 b:{*.a.{..{f(\da)}.{g(\da)}}}
\whichgives
\formula p8|{..{..{*-a.{..{f(\da)}.{g(\da)}}}
                 -{f(b)}}
              .{g(b)}}
\named "59"
\endverbatim\Eend1


% %%%%%%%%%%%
\beginsection 9. Terminal symbols of the Begriffsschrift

In part~3 of the {\it Begriffsschrift\/} four symbols are
defined. They are available as \TeX\ commands. The commands get more
flexibility than the {\it Begriffsschrift\/} requires as all possible
content can be addressed by parameters.

\item{1.}¤\gfvererbe¤ with four parameters:
¤$\gfvererbe\kd\ka{x(\ka)}{f(\kd,\ka)}$¤ outputs
$\gfvererbe\kd\ka{x(\ka)}{f(\kd,\ka)}$.
\item{2.}¤\gffolgt¤ with three parameters:
¤$\gffolgt\kg\kb{f(x_\kg,y_\kb)}$¤ outputs $\gffolgt\kg\kb{f(x_\kg,y_\kb)}$.
\item{3.}¤\gfgehoertan¤ with three parameters:
¤$\gfgehoertan\kg\kb{f(x_\kg,z_\kb)}$¤ outputs
$\gfgehoertan\kg\kb{f(x_\kg,z_\kb)}$.
\item{4.}¤\gfeindeutig¤ with three parameters:
¤$\gfeindeutig\kd\ke{f(\kd,\ke)}$¤ outputs $\gfeindeutig\kd\ke{f(\kd,\ke)}$.

\noindent
I use shortcuts for these special terminal symbols and reduce the
number of parameters to the amount that is used in the text;
see Example~11 below.


% %%%%%%%%%%%
\beginsection 10. Output style of the Grundgesetze

\toggleGGstyle
Frege made changes to the notation in his main work. A command and
several flags are defined to switch from the style of the {\it
Begriffsschrift\/} to the style of the {\it Grundgesetze}.
\smallskip
\item{1.}¤\toggleGGstyle¤ switches the styles; the default is the style of
the {\it Begriffsschrift}. Several of the changes are global so even
if the switch is made inside a group it shall be called a second time
to switch it back. This happens when it is called for the first time:
\itemitem{a)}¤\gfbsuseGGstyletrue¤ is set. Together with a call to
¤\gfbsneuestriche¤ the thickness of the lines is changed.
\itemitem{b)}¤\gfbskompakttrue¤ is set. This reduces the width of the formulas
that are input with the short form.
\itemitem{c)}¤\gfggschlussweisetrue¤ is set. This activates more symbols
to express the eight inferences of the {\it Grundgesetze\/} (see Section~12).
\itemitem{d)}¤\gfbsfonoohnepunkttrue¤ is set. No period after the number of
a formula is set in ¤\fono¤ or ¤\named¤.
\itemitem{e)}¤\gfbsnegdirekttrue¤ is set. The negation indicator touches
the content stroke.
\item{2.}The flags can be set individually, for example, to get a compact output
but keep the line thicknesses etc.\ of the {\it Begriffsschrift}.
\item{3.}To control the compact output of the generality when the flag
¤\gfbskompakttrue¤ is set more flags are available. Examples are given to
show how these flags influences the output.
\itemitem{$\alpha$)}The output is not compact if ¤\gfbskeinekompaktehoehlungtrue¤
is set. At the left the flag is false and at the right true; the source code
is ¤\frege|{..{*-a.{\gY(\da)}}.{*.e.{\gF(\de)}}}¤:
\smallskip\indent\indent
{\gfbskompakttrue \gfbskeinekompaktehoehlungfalse
\frege|{..{*-a.{\gY(\da)}}.{*.e.{\gF(\de)}}}
}\quad
{\gfbskompakttrue \gfbskeinekompaktehoehlungtrue
\frege|{..{*-a.{\gY(\da)}}.{*.e.{\gF(\de)}}}
}.
\itemitem{$\beta$)}The output of the short form rule~1 (with the {\tt=}) can
have no effect except if the flag named ¤\gfbsaussagesichtbartrue¤ is set.
At the left the flag is false and at the right true; the line reads
¤\frege.{=.a} vs \frege.{=.{=.a}}¤:
{\gfbskompakttrue \gfbsaussagesichtbarfalse
\frege.{=.a} vs \frege.{=.{=.a}}
}\quad
{\gfbskompakttrue \gfbsaussagesichtbartrue
\frege.{=.a} vs \frege.{=.{=.a}}
}

{\it Note\/}: Both input forms, the symbolic representation and the
short form, can be used with this style. But only the short form
creates smaller formulas if ¤\gfbscompakttrue¤ is set. The width of a
formula in symbolic representation must be changed manually if it is
not coded with the symbols of the last list in Section~5.

\Example8
Here is the example of Sections 1 and~5; it is coded in the
short form as ¤\frege|{..\gA.{*.a.{f(\da)}}}¤:
\frege|{..\gA.{*.a.{f(\da)}}}.
\Eend0


% %%%%%%%%%%%
\beginsection 11. Setting formulas in parentheses and brackets (Grundgesetze)

In the {\it Grundgesetze\/} some formulas must be placed in
parentheses or brackets. A few commands are provided to make this an
easy task.
\smallskip
\item{1.}¤\pfrege<x><j>{<f>}¤ acts like ¤\frege<j>{<f>}¤ and adds parentheses around
this formula. The result is stored in a box with the number {\tt<x>}. In total
there are four boxes for this task.
\item{2.}¤\bfrege<y><j>{<f>}¤ is like ¤\pfrege¤ but the formula is placed between
brackets. The pool of boxes is independent from the ¤\pfrege¤ boxes and there are
four boxes.
\item{3.}¤\pparens<x><M>¤ places the math {\tt<M>} with parentheses in the box
numbered {\tt<x>}. The boxes are taken from the pool that is available for
¤\pfrege¤.
\item{4.}¤\bparens<y><M>¤ acts like ¤\pparens¤ but it uses brackets and the
pool of boxes of ¤\bfrege¤.
\item{5.}¤\pfbox<x>¤ outputs box number {\tt<x>} from the ¤\pfrege¤ pool.
\item{6.}¤\bfbox<y>¤ outputs the box {\tt<y>} from the pool used by ¤\bfrege¤.

\Example9
The following formula appears on page~115 of the {\it Grundgesetze},
vol.~1.
\Epause
\smallskip
\begingroup
\pfrege1.{.-{\ke=c}.{\ke\G^v}}
\pfrege2.{.-{\ke=b}.{\ke\G^u}}
\bfrege1.{-.{\ke=n}.{{\ke\G^\kesl\pfbox1}}}
\bfrege2.{-.{\ke=m}.{{\ke\G^\kesl\pfbox2}}}
\bfrege3.{--{{\ke\G^(\ka\G^q)}}-{..{\ke=b}-{\ka=c}}}
\pfrege1.{-.{\ke=n}.{{\ke\G^\kesl\pfbox1}}}
\pfrege2.{-.{\ke=m}.{{\ke\G^\kesl\pfbox2}}}
\pparens3{\kesl\bfbox1\G^\G>q}
\pparens4{\bfbox1\G^\G>\kasl\kesl\bfbox3}

$$\frege|{..{..{..{{\kesl\pfbox2\G^\pfbox4}}
                 -{{b\G^\kesl\pfbox2}}}
              -{{c\G^\kesl\pfbox1}}}
           .{{\kesl\pfbox2\G^\pfbox3}}}
$$
\endgroup
\Econt
And here is the source:
\verbatim
\pfrege1.{.-{\ke=c}.{\ke\G^v}}
\pfrege2.{.-{\ke=b}.{\ke\G^u}}
\bfrege1.{-.{\ke=n}.{{\ke\G^\kesl\pfbox1}}}
\bfrege2.{-.{\ke=m}.{{\ke\G^\kesl\pfbox2}}}
\bfrege3.{--{{\ke\G^(\ka\G^q)}}-{..{\ke=b}-{\ka=c}}}
\pfrege1.{-.{\ke=n}.{{\ke\G^\kesl\pfbox1}}}
\pfrege2.{-.{\ke=m}.{{\ke\G^\kesl\pfbox2}}}
\pparens3{\kesl\bfbox1\G^\G>q}
\pparens4{\bfbox1\G^\G>\kasl\kesl\bfbox3}
\endverbatim\smallskip\verbatim
$$\frege|{..{..{..{{\kesl\pfbox2\G^\pfbox4}}
                 -{{b\G^\kesl\pfbox2}}}
              -{{c\G^\kesl\pfbox1}}}
           .{{\kesl\pfbox2\G^\pfbox3}}}
$$\endverbatim\Eend1


% %%%%%%%%%%%
\beginsection 12. Inferences of the Grundgesetze

The list entry~1c of Section~10 needs some explanation as it extends
the list of available symbols for inferences. Two of the symbols have
no output ({\tt?}\ and {\tt>}). In the following list the text in
parentheses contains the sequence ¤\*?#"¤ or ¤\*?>#¤ for the symbol
`{\tt\#}'.
\smallskip
\item{\tt ?} (must be the first in the triple) signals that the next
two symbols build an inference line in the style of the {\it
Grundgesetze\/}. Such symbols use 50\% more space than ordinary symbols.
\item{\tt -}(\*?-") is a single horizontal stroke;
\item{\tt .}(\*?.") is a single (centered) period;
\item{\tt =}(\*?=") is a double stroke;
\item{\tt *}(\*?*") is a single stroke in the height of the upper line of the
double stroke;
\item{\tt\char95 }(\*?_") is an empty space;
\item{\tt "}(\*?"") is a skip;
\item{\tt >}has no output; the next symbol must be a {\sl transition-sign\/}:
\itemitem{\tt x}(\*?>x) it must follow a {\tt >};
\itemitem{\tt u}(\*?>u) it must follow a {\tt >}.

\noindent
Inferences of the {\it Grundgesetze\/} are separated by a new type of line:
\item{1.}¤\separator¤ outputs a thick line with a centered dot.

\Example{10}
Here is an inference chain from the {\it Grundgesetze\/}:
\smallskip
\advance\rightskip by 2\rightskip
\use p3"III"
\formula p5|{.-{*.f.{..{\df(a)}.{\df(a)}}}-{a=a}}
\nlp{6.5}\*?>x
\formula p5|{..{a=a}.{*.f.{..{\df(a)}.{\df(a)}}}}
\named "$\ka$"
\separator
\use p3"I"
\formula p5|{..{f(a)}.{f(a)}}
\nlp5\*?_"\*?>u
\formula p5|{*.f.{..{\df(a)}.{\df(a)}}}
\named"$\kb$"
\bcc3/m$\ka$:.\*__"\rep3\*?--\ecc
\formula p5|{=.{a=a}}
\named"IIIe"

\rightskip=\leftskip
\medskip
And this is the code:
\verbatim
\use p3"III"
\formula p5|{.-{*.f.{..{\df(a)}.{\df(a)}}}-{a=a}}
\nlp{6.5}\*?>x
\formula p5|{..{a=a}.{*.f.{..{\df(a)}.{\df(a)}}}}
\named "$\ka$"
\separator
\use p3"I"
\formula p5|{..{f(a)}.{f(a)}}
\nlp5\*?_"\*?>u
\formula p5|{*.f.{..{\df(a)}.{\df(a)}}}
\named"$\kb$"
\bcc3/m$\ka$:.\*__"\rep3\*?--\ecc
\formula p5|{=.{a=a}}
\named"IIIe"
\endverbatim\Eend1


% %%%%%%%%%%%
\beginsection 13. Terminal symbols of the Grundgesetze

The {\it Grundgesetze\/} has not such complex constructions as the
{\it Begriffsschrift\/} (see Section~9). But it contains a lot of
special symbols.
\smallskip
\settabs 2 \columns
\def\hsnor#1.{\hbox to 3em{\hss#1.}\ }
\+\hsnor1.¤\G!¤: $\G!\kx$.&
\hsnor16.¤\Gfm¤: $u\Gfm q$.\cr
\+\hsnor2.¤\kasl¤: $\kasl\gF(\ka)$.&
\hsnor17.¤\Gsm¤: $p\Gsm q$.\cr
\+\hsnor3.¤\G^¤: $a\G^b$.&
\hsnor18.¤\G<¤: $p\G<q$.\cr
\+\hsnor4.¤\GI¤: $\GI p$.&
\hsnor19.¤\Gy¤: $p\Gy q$.\cr
\+\hsnor5.¤\G>¤: $\G>p$.&
\hsnor20.¤\Ge¤: $p\Ge q$.\cr
\+\hsnor6.¤\GU¤: $\GUorg p$ (or $\GU p$).&
\hsnor21.¤\Gc¤: $p\Gc q$.\cr
\+\hsnor7.¤\Gn¤: $\Gnorg u$ (or $\Gn u$).&
\hsnor22.¤\Gmf¤: $p\Gmf q$.\cr
\+\hsnor8.¤\G0¤: $\G0$.&
\hsnor23.¤\G*¤: $\G*t$.\cr
\+\hsnor9.¤\G1¤: $\G1$.&
\hsnor24.¤\G6¤: $\G6s$.\cr
\+\hsnor10.¤\Gbf¤ or ¤\Gff¤: $\Gbf$.&
\hsnor25.¤\Gf¤: $\Gf s$.\cr
\+\hsnor11.¤\G-¤: $\G-q$.&
\hsnor26.¤\GF¤: $s\GF u$.\cr
\+\hsnor12.¤\G~¤: $\G~q$.&
\hsnor27.¤\Gl¤: $s\Gl u$.\cr
\+\hsnor13.¤\G_¤: $p\G_q$.&
\hsnor28.¤\Gp¤: $\Gp s$.\cr
\+\hsnor14.¤\Guu¤: $\Guu$.&
\hsnor29.¤\Gss¤ or ¤\Gsz¤: $s\Gss p$.\cr
\+\hsnor15.¤\G;¤: $o\G;a$.\cr

\noindent
The symbols are constructed with the Computer Modern
and \AmSTeX\ fonts.  But with the external flag ¤\fgefontsknown¤ the
second symbol in numbers 6 and 7 is selected (see Section~3).

Several symbols are rotated and this requires the package {\tt
rotate.tex}. The output of unrotated symbols is activated by default;
the external flag ¤\let\showrotation=t¤ must be given to activate
rotation (see Section~3).
\toggleGGstyle


% %%%%%%%%%%%
\beginsection 14. Typesetting definitions

The following command is provided to typeset definitions:
\smallskip
\item{1.}¤\gfdefinition p<n><J> <j>{<f>}=<e>¤ outputs the formula
¤\frege<j>{<f>}¤ in parentheses and sets it equal to the new symbol
{\tt<e>} if the parameter {\tt<J>} is either {\tt|} to type a judgment
or {\tt.}\ for a definition.  In the style of the {\it Grundgesetze\/}
the parameter {\tt<J>} can also be {\tt:}. In this case write ¤<j><f>¤
and this is interpreted as a math formula and not as a formula created
by ¤\frege¤.  The parameter {\tt<n>} is the position parameter to move
the output the given amount of units to the right.

\noindent
The output of a definition depends on the style.

\Example{11}
The definition in formula~69 of the {\it Begriffsschrift\/} is:
\smallskip
\def\1#1{{\gfvererbe\kd\ka{#1(\ka)}{f(\kd,\ka)}}}
\gfdefinition p4. .{*.d.{..{*.a.{..{F(\da)}.{{f(\dd,\da)}}}}.{F(\dd)}}}={\1F}
\smallskip
In the style of the {\it Grundgesetze\/} the output is:
\smallskip
\toggleGGstyle
\def\1#1{{\gfvererbe\kd\ka{#1(\ka)}{f(\kd,\ka)}}}
\gfdefinition p4. .{*.d.{..{*.a.{..{F(\da)}
                                  .{{f(\dd,\da)}}}}
                          .{F(\dd)}}}={\1F}
\toggleGGstyle
\smallskip
And here is the source (note that ¤\1¤ is a normal macro and needs no
macro expansion):
\verbatim
\def\1#1{{\gfvererbe\kd\ka{#1(\ka)}{f(\kd,\ka)}}}
\gfdefinition p4. .{*.d.{..{*.a.{..{F(\da)}
                                  .{{f(\dd,\da)}}}}
                          .{F(\dd)}}}={\1F}
\endverbatim\Eend1

\Example{12}
Here is an example of the {\it Grundgesetze}, in which the {\tt<J>} is
a colon.
\smallskip
\advance\rightskip by 2\rightskip
\toggleGGstyle
\gfdefinition p4: \kesl(o\G^(a\G^\ke))={o\G;a}
\named "$\gX$"
\toggleGGstyle
\smallskip
\rightskip=\leftskip
And this is the source:
\verbatim
\gfdefinition p4: \kesl(o\G^(a\G^\ke))={o\G;a}
\named "$\gX$"
\endverbatim\Eend1


% %%%%%%%%%%%
\beginsection 15. Commands to improve page breaks

The formulas are never broken. A page or column break must appear
at a decollator, separator, or after an inference rule. To gain some
space the formulas can be typeset in a smaller height and depth.
(¤\offinterlineskip¤ should be set.)
\smallskip
\item{1.}¤\gfbssetzeabstand<h><d>¤ sets the height and depth between the lines to
{\tt<h>}\thinspace pt and {\tt<d>}\thinspace pt, resp.;
\item{2.}¤\gfbsreduziereabstandum<nn>¤ reduces the height and depth each by {\tt<nn>}\%
({\tt<nn>} is a two-digit number);
\item{3.}¤\gfbsabstandzuruecksetzen¤ sets height and depth to the values of ¤\strutbox¤.

\Example{13}
The right formula uses ¤\gfbsreduziereabstandum20¤:
\smallskip
\begingroup
\offinterlineskip
\setbox0=\hbox to .3\hsize{\vtop to 30pt{%
               \nlp4\*_[-\*-:-\*---\ce{$a$}
               \nlp4\*___\*_'-\*-:-\ce{$c$}
               \nlp4\*___\*___\*_'-\ce{$e$}\vss}\hss}
\setbox1=\hbox to .3\hsize{\vtop to 30pt{%
               \nlp1\*_[-\*-:-\*---\ce{$a$}
               \gfbsreduziereabstandum20
               \nlp1\*___\*_'-\*-:-\ce{$c$}
               \nlp1\*___\*___\*_'-\ce{$e$}\vss}\hss}
\centerline{\box0 \box1 }
\gfbsabstandzuruecksetzen
\endgroup

And this is the code:
\verbatim
\offinterlineskip
\setbox0=\hbox to .3\hsize{\vtop to 30pt{%
               \nlp4\*_[-\*-:-\*---\ce{$a$}
               \nlp4\*___\*_'-\*-:-\ce{$c$}
               \nlp4\*___\*___\*_'-\ce{$e$}\vss}\hss}
\setbox1=\hbox to .3\hsize{\vtop to 30pt{%
               \nlp1\*_[-\*-:-\*---\ce{$a$}
               \gfbsreduziereabstandum20
               \nlp1\*___\*_'-\*-:-\ce{$c$}
               \nlp1\*___\*___\*_'-\ce{$e$}\vss}\hss}
\centerline{\box0 \box1 }
\gfbsabstandzuruecksetzen
\endverbatim\Eend1


% %%%%%%%%%%%
\beginsection 16. List of parameters

Nearly all aspects of the lines that build the structure of Frege's
notation is controlled by parameters. They can be changed to fit
better with a certain type of font. But note that a couple of tests
have been made with these values to guarantee, for example, that the
affirmation can be placed in front of a quantification.\\

\noindent
Here is the list of registers that make up the style of the
{\it Begriffsschrift}.
\item{1.}$¤\gfbsvolleeinheit¤=\rm0.57\,em$; width of one ¤\*¤ triple
\item{2.}$¤\gfbsraise¤=\rm0.5\,ex$; height of the horizontal line
\item{3.}$¤\gfbsneg¤=\rm0.25\,ex$; height of the negation indicator
\item{4.}$¤\gfbsstrichdicke¤=\rm0.5\,pt$; thickness of the horizontal line
\item{5.}$¤\gfbsurteildicke¤=\rm1\,pt$; thickness of the judgment stroke\\
\item{6.}$¤\gfbsuht¤=\rm1.5\,ex$; height of the judgment stroke\\
\item{7.}$¤\gfbsudp¤=\rm0.5\,ex$; depth of the judgment stroke\\
\item{8.}$¤\gfbsersetzungdicke¤=\rm0.8\,pt$; thickness of the substitution line\\
\item{9.}$¤\gfbsdefdicke¤=\rm0.75\,pt$; thickness of the definition strokes
\item{10.}$¤\gfbsschlussabstand¤=\rm2.5\,pt$; distance of double lines in an inference
\item{11.}$¤\gfbsschlussdicke¤=\rm0.8\,pt$; thickness of the second line in an inference
\item{12.}$¤\gfbsdpabstand¤=\rm0.2\,em$; distance between colons in front of an inference
\smallskip
\item{13.}$¤\gfbsmaxanzahlzeilen¤=25$; the maximal amount of lines
in a single formula of the short form (the package declares for each
line one or two token and skip registers; two registers of each kind
are needed if substitutions are allowed);
\item{14.}$¤\gfbsnegpct¤=70$; when ¤\ifgfbsnegdirect¤ is set to false the negation
indicator has only a height of this percentage so that a gap occurs.
\smallskip
\noindent
And this is the list of additional dimen parameters for the style of
the {\it Grundgesetze}.
\item{15.}$¤\gfggstrichdicke¤=\rm0.58\,pt$; common line thickness
\item{16.}$¤\gfggraise¤=\rm0.14\,ex$; height of horizontal line
\item{17.}$¤\gfggneg¤=\rm0.47\,ex$; height of negation indicator
\item{18.}$¤\gfgguht¤=\rm1.4\,ex$; height of the judgment stroke
\item{19.}$¤\gfggudp¤=\rm0.9\,ex$; depth of the judgment stroke
\item{20.}$¤\gfggdpabstand¤=\rm0.06\,em$; distance between colons in front of an inference


% %%%%%%%%%%%
\beginsection 17. Related work

Other people have worked on the problem to typeset Frege's notation
in \TeX\ too. J.~Parsons has developed a package {\it begriff.sty\/}
for \LaTeX\ (see [7]). Later the package was improved by
Q.~Pamp~([8]). A team that translated the {\it Grundgesetze\/} created
its own package from {\it begriff.sty\/} and called it {\it
grundgesetze.sty\/} (see [9]). It uses the fonts [5] that J.~J.~Green
has developed for the special terminal symbols in the {\it
Grundgesetze}.


% %%%%%%%%%%%
\beginsection 18. References

\raggedright\frenchspacing
\item{[1]}Gottlob Frege,
   {\sl Begriffsschrift, eine der arithmetischen nachgebildete Formelsprache
   des reinen\hfil\break Denkens\/}
   (Halle an der Saale: Louis Nebert, 1879)
\item{}``Begriffsschrift, a formula language, modeled upon that of arthmetic, for
   pure thought,'' in: Jean van Heijenoort (ed.),
   {\sl From Frege to G\"odel: A Sourcebook in Mathematical Logic,
   1879--1931\/},\hfil\break(Cambridge, MA: Harvard University Press, 1967), 1--82;
   translation by S. Bauer-Mengelberg    

\item{[2]}Gottlob Frege,
   {\sl Grundgesetze der Arithmetik --- begriffsschriftlich abgeleitet\/}
   (Jena: Hermann Pohle, Volume 1 1893, Volume 2 1903)
\item{}{\sl Basic Laws of Arithmetic\/} (Oxford: Oxford Univ. Press, 2013);
   translation by Philip A. Ebert, Marcus Rossberg, Crispin Wright
   
\item{[3]}American Mathematical Society, {\tt amstex},
   ¤http://www.ams.org/publications/authors/tex/tex¤ or
   ¤http://www.ctan.org/tex-archive/macros/amstex¤
   (accessed: 2014-11-29)

\item{[4]}{\tt rotate},
   ¤http://www.ctan.org/tex-archive/macros/plain/contrib/misc¤
   (accessed: 2014-11-29)

\item{[5]}J. J. Green, {\tt fge},
   ¤http://www.ctan.org/tex-archive/fonts/fge¤
   (accessed: 2014-11-29)

\item{[6]}Silvio Levy, ``Using Greek fonts with \TeX,''
   TUGboat {\bf 9,1} (1988), 20--24

\item{[7]}Josh Parsons, {\tt begriff.sty},
   ¤http://www.ctan.org/tex-archive/macros/latex/contrib/begriff¤
   (accessed: 2014-11-29)

\item{[8]}Quirin Pamp, {\tt frege.sty},
   ¤http://www.ctan.org/tex-archive/macros/latex/contrib/frege¤
   (accessed: 2014-11-29)

\item{[9]}Marcus Rossberg, {\tt grundgesetze.sty},
   ¤http://www.ctan.org/tex-archive/macros/latex/contrib/¤\hfil\break¤grundgesetze¤
   (accessed: 2014-11-29)

\bye
% Local Variables:
% coding: iso-latin-9-unix
% End:
