% This file is public domain.
% Originally written 1992, Don Hosek.
% This declaration added by Clea F. Rees 2008/11/16 with the permission of Dan Hosek.
%
\input resumemac
\heading{center}
\name{Donald A. Hosek}
\address{\bf Home:\\
	3916 Elmwood Avenue\\
	Stickney, IL 60402}
\address{\bf School:\\
	Platt Campus Center\\
	Harvey Mudd College\\
	Claremont, CA 91711}
\tag{Job Objective}
Employment for the period from May, 1987 to August, 1987 doing technical writing
and/or programming
\endtag
\tag{Previous Employment}
Jobs held were:
\(Nov.~1986--Present)\TeX\ consultant, Harvey Mudd College Mathematics
Department. Duties: answering questions about the \TeX\ computer typesetting
system; writing special macros; and creating documentation.\\
\(June~1986--Aug.~1986)Word processing consultant, University of Illinois at
Chicago. Duties: Answering questions about the word processing packages
supported by the University of Illinois Computer Center (SCRIPT, GML, \TeX);
and writing macro packages for \TeX.\\
\(Jan.~1986--Aug.~1986)General Consultant, University of Illinois at Chicago.
Duties: Answering questions from users about the computer center's IBM mainframe
running VM/CMS and MVS---support included FORTRAN, PL/1, Pascal, 370 Assembler,
C, and JCL; tending the Xerox~8700, IBM~3270, and General Electric printers; and
programming projects to augment system software.\\
\endtag
\tag{Education}
Eventual educational goal is to have doctorates in English and mathematics.
\(Sep.~1986--Present)Harvey Mudd College. Currently working on B.~S. degree in
a double major of mathematics and English.\\
\(Sep.~1985--Aug.~1986)University of Illinois at Chicago. Additional studies.
Coursework consisted of mathematics and humanities classes.\\
\endtag
\tag{Additional work}
Since January, 1987, I have been publishing an electronic ``magazine'',
\TeX MaG, distributed through the BITNET and InterNET electronic networks. The
magazine is intended to distribute macros, product information, etc., to users
of the \TeX\ computer typeseting system. It is distributed to nearly two hundred
people around the world.

Since September, 1984, I have been involved in student journalism. My
involvement has included photography (taking, developing, and printing
pictures), and writing (including a bi-weekly column written for a year).
\endtag
\tag{Hobbies}
Music (I play bass, piano, guitar, and sing), writing, and photography.
\endtag
\bye
	G'Day
