%%
%% This is the documentation of DPcircling package.
%% (Last Update: 2020/04/15)
%% Maintained on GitHub:
%% https://github.com/domperor/DPcircling
%%
%% Copyright (c) 2020 Oura M. (domperor)
%% Released under the MIT license
%% https://opensource.org/licenses/mit-license.php
%% 

\documentclass[10pt]{article}
\usepackage[dvipdfmx]{graphicx}
\usepackage{geometry}
\usepackage{url}
\geometry{a4paper,total={170mm,257mm},left=20mm,top=20mm}
\usepackage{DPcircling}
\title{DPcircling package v1.0}
\author{Oura M. (domperor)}
\date{\today}
\pagestyle{empty}
\begin{document}
\maketitle\thispagestyle{empty}

\section*{about this package}

This package provides 4 types of text decorations: \verb+\DPcircling+ \DPcircling{circle}, \verb+\DPrectangle+ \DPrectangle{rectangle}, \verb+\DPjagged+ \DPjagged{jagged rectangle}, and \verb+\DPfanshape+ \DPfanshape{fan-shape}. The baseline would be adjusted properly according to the surroundings\footnote{Unless you cram too tall things into a small decoration box. The box would be broken, and you would make a mess.}. You can use these decorations both in text mode and in math mode. You can specify \verb+line color+, \verb+line width+, \verb+width+, and \verb+height+ as option keys.

\ 

This package is maintained on GitHub: \url{https://github.com/domperor/DPcircling}

\subsection*{basic usage}

\begin{quote}
(in preamble): \verb+\usepackage[+\textit{\texttt{driver}}\verb+]{graphicx}\usepackage{DPcircling}+\footnote{The author personally uses dvipdfmx as the driver.}

(in your text): \verb+\DP*****[+\textit{\texttt{options}}\verb+]{+\textit{\texttt{content}}\verb+}+
\end{quote}

For instance, the code
\begin{quote}
\begin{verbatim}
It is a \DPjagged[line color=blue,line width=1.4pt]{\color{brown}great} opportunity.
\end{verbatim}
\end{quote}
\noindent gives the result:
\begin{quote}
It is a \DPjagged[line color=blue, line width= 1.4pt]{\color{brown}great} opportunity.
\end{quote}

\subsection*{required packages}

DPcircling requires the following packages: \verb+tikz+, \verb+keyval+, \verb+graphicx+, and the ones that these packages require.

\subsection*{aliases}

\verb+\DPcircle+ and \verb+\DPcirc+ are the aliases of \verb+\DPcircling+.
\verb+\DPrect+ is the alias of \verb+\DPrectangle+.

\subsection*{changing default values}

The default values of \verb+line color+, \verb+line width+, \verb+width+, and \verb+height+ are \verb+black+, \verb+1pt+, noted below(*), and \verb+2*(content height)+. You can modify these like this:
\begin{quote}
\begin{verbatim}
\DPcirclingDefault{line color=brown, line width=0.33pt, width=4em, height=5em}
\end{verbatim}
\end{quote}

\noindent (*) The default values of \verb+width+ are \verb+max{2*(content width), 2em}+ (circle) and 
\verb:(content width)+2em: (else).

\section*{version history}

2020/04/15 v1.0

\end{document}